---
title: "p8105_hw2_nja2140"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
```

Problem 1: Read and Clean the Mr. Trash Wheel dataset.
  -Specify the sheet in the Excel file and to omit non-data entries (rows with notes /     figures; columns containing notes) using arguments in read_excel
  -Use reasonable variable names
  -Omit rows that do not include dumpster-specific data
  -Round the number of sports balls to the nearest integer and converts the result to an    integer variable (using as.integer)
  
```{r}
trash_data = 
  read_excel("./Trash_Wheel.xlsx", sheet = 1) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  select(-15,-16,-17) %>%
  mutate(sports_balls = as.integer (sports_balls))
trash_data
```
Read and clean precipitation data for 2017. Omit rows without precipitation data and add a variable year.

```{r}
trash17_data=
  read_excel("./Trash Wheel.xlsx", sheet = 6, skip = 1) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = 2017)
```

Read and clean precipitation data for 2018. Omit rows without precipitation data and add a variable year.

```{r}
trash18_data=
  read_excel("./Trash Wheel.xlsx", sheet = 5, skip = 1) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = 2018)
trash18_data
```

Combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful

```{r}
precip_data=
  full_join(trash17_data, trash18_data) %>%
  mutate(month = month.abb[month]) 
precip_data
```

Problem 2: Clean the pols-month dataset. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable

```{r}
pol_data = 
  read_csv("./FiveThirtyEight/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, c("year", "month", "day")) %>%
  mutate(
    year = as.numeric(year),
    month = as.numeric(month),
    month = month.name[month]
    ) %>%
  mutate(president = ifelse(prez_gop ==1, "gop", "dem")) %>%
  select(-prez_gop, -prez_dem, -day)
pol_data
```
Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
snp_data =
  read_csv("./FiveThirtyEight/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, c("month", "day", "year")) %>%
  mutate(year = as.numeric(year),
    month = as.numeric(month),
    month = month.name[month]
    ) %>%
  select(year, month, -day, close) %>%
  arrange (year, month)
snp_data
```
Tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r}
une_data = 
  read_csv("./FiveThirtyEight/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "rate"
  )
une_data
```
Join the datasets by merging snp into pols, and merging unemployment into the result
```{r}
snppols_data=
  full_join(snp_data, pol_data)
total_data=
  inner_join(snppols_data, une_data, by = "year")
total_data
```
