---
title: "p8105_hw2_nja2140"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Problem 1: Read and Clean the Mr. Trash Wheel dataset.
  -Specify the sheet in the Excel file and to omit non-data entries (rows with notes /     figures; columns containing notes) using arguments in read_excel
  -Use reasonable variable names
  -Omit rows that do not include dumpster-specific data
  -Round the number of sports balls to the nearest integer and converts the result to an    integer variable (using as.integer)
```{r}
library(readxl)
trash_data = 
  read_excel ("./Trash Wheel.xlsx", sheet = 1) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  select(-15,-16,-17) %>%
  mutate(sports_balls = as.integer (sports_balls))
trash_data
```
Read and clean precipitation data for 2017. Omit rows without precipitation data and add a variable year.

```{r}
trash17_data=
  read_excel("./Trash Wheel.xlsx", sheet = 6, skip = 1) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = 2017)
```
Read and clean precipitation data for 2018. Omit rows without precipitation data and add a variable year.

```{r}
trash18_data=
  read_excel("./Trash Wheel.xlsx", sheet = 5, skip = 1) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(year = 2018)
trash18_data
```

Combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful

```{r}
precip_data=
  full_join(trash17_data, trash18_data) %>%
  mutate(month = as.character (month)) 
precip_data
```

